Ornaments as indicators of social changes and cultural practices in northeastern Taiwan before and after European colonial period

### 2019-12-11 analysis/paper/paper.Rmd

The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BC) and a lower component (700-1200 AD, 1200-800 BC) separated by a sterile deposit spanning c. 150 years. These component divisions are based on the differences in the colour and texture of the deposit. The interpration of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007]. There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data). We focus on the upper component because only this componant spans the periods of pre-European contact, European presence, and Chinese presence. In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases. Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011]. However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artefacts to the pre-European, European, and the Chinese periods.

>>>>

The chronology of Kiwulan can be divided into two phases represented by a upper component (1400-1900 AD, 600-100 BP) and a lower component (700-1200 AD, 1200-800 BP) separated by a sterile deposit spanning c. 150 years. These component divisions are based on the differences in the colour and texture of the deposit. The interpration of the sterile deposit is still under debate, with pollen analysis suggesting dry weather leading to site abandonment [@Lin2015; @Chiu2004; @Chen2007]. There are 32 radiocarbon ages spanning the two components, previously published by @Chen2007, and shown here in Figure \@ref(fig:C14-data) and Table \@ref(tab:C14-table-data). We focus on the upper component because only this componant spans the periods of pre-European contact, European presence, and Chinese presence. In the upper component, all excavation squares in our sampling area show signs of continuous human occupation during each of the three phases. Previous work divided the upper component into six analytical units, spanning from the 14th century to the 19th century, according to the radiocarbon dates, excavation depth, consistency of contexts, and types of chronologically diagnostic ceramics such as blue and white porcelains [@Hsieh2009; @Wang2011]. However, we found some ambiguities in the previous chronology, so to help answer our research question, we re-examined the upper component to devise a new chronology to assign artefacts to the pre-European, European, and the Chinese periods.
